---
title: "Pella-Tomlinson Model Analysis"
author: "Matthew Holden"
date: "28 July 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Recruitment functions
```{r functions}
### Pella-Tomlinson Surplus Production function 
  # (phi controls strength of weak allee effect phi=1 recovers logistic/schaeffer)
  PT = function(s, pars){
    r = pars[1]; k = pars[2]; phi = pars[3];
    y = s + r* (phi/(phi + 1)) *s*(1 - (s/k)^phi)
    return(y)
  }
```

## Parameters and Data

```{r parameters}
  #parameters
    rho = 1       #discrete discount factor
    r = 0.25      #growth rate
    k = 100       #carying capacity
    sig = .4      #std of normal rv (for lognormal rvs)
    m = -.5*sig^2 #mean of normal rv such that lognomormal has mean 1
    phi = 10         #power in logistic function
    N=200
    parms = c(r,k,phi)
    
    #data (x = escapement, y = recruitment, z = random variable)
    x = runif(N,0,1.02*k)
    z = rlnorm(N,meanlog = m, sdlog = sig)
    #z = runif(N, .4,1.6) #use this for uniform data
    y = z*PT(x,parms)
    x1 = seq(0,k,length.out=N)
    y1 = PT(x1,parms)
  
  #plot data
    plot(x,y)
    lines(x1, y1)
```


## Optimal Escapement
```{r Optimal}
  OptEscPT = k/r*(sqrt(rho*(1+r)) -1)
  OptEscL = k/(1+r)
  OptEscL = .5*k
  OptEscPTp = (1/(phi+1))^(1/phi) * k
```

Logistic escapement is `r OptEscL` and Paella Tomilson Escapement is $\left[\frac{1}{\phi+1}\right]^{1/\phi}k$, with $\phi=$ `r phi` escapement is `r OptEscPTp`. Increasing $\phi$ increases optimal escapment

```{r PaellaTomilson, echo=FALSE}

##plot optimal escapement as a function of exponent in pella - T model
  phiV=seq(1,70)
  OptEscPTV = (1/(phiV+1))^(1/phiV) * k
  plot(phiV, OptEscPTV, xlab='phi', ylab='Optimal Escapement', main='Paella Tomilson Optimal Escapement, k=100')
```

## Data fitting


```{r SumSquaredErrorFunctions}
  
  ssePT = function(p,x,y){
    yMod = PT(x,p)
    return( sum( (y-yMod)^2 ) )
  }

  sseL = function(p,x,y){
    yMod = PT(x,c(p,1))
    return( sum( (y-yMod)^2 ) )
  }
  
  parInit = c(r,k)
  optOutPT = optim(c(parInit,phi), fn = ssePT, x=x, y=y)
  optOutL = optim(parInit, fn = sseL, x=x, y=y)
  
  #plot fit
    plot(x,y)
    xSort = sort(x)
    lines( xSort, PT(xSort, optOutPT$par ) )
    lines( xSort, PT(xSort, c(optOutL$par,1) ), col ='red')
  

### Calculate Optimal Escapements

    
  rPT = optOutPT$par[1];
  kPT = optOutPT$par[2]; 
  phiPT = optOutPT$par[3];
  
  rL = optOutL$par[1];
  kL = optOutL$par[2]; 
  
  OptEscPT = kPT*(1/(1 + phiPT))^(1/phiPT)
  OptEscL = .5*kL
```
Optimal escapement in the fitted PT model is `r OptEscPT` and fitted Logistic is `r OptEscL`


## Test Escapement Strategies
```{r objective}
## calculate the sum of harvest for fitted PT model    
  xPT=rep(NA,N+1)
  xPT[1]=k    
  objPT=0    
  for(i in 1:N){
    xPT[i+1] = PT(xPT[i], c(r, k, phi))
    h = max(0, xPT[i+1] - OptEscPT)
    xPT[i+1] = xPT[i+1] - h
    objPT = objPT + h
  }    
  ObjPT = objPT + xPT[N+1] #take all fish at eNd

## calculate the sum of harvest for fitted L model   
  xL=rep(NA,N+1)
  xL[1]=k    
  objL=0    
  for(i in 1:N){
    xL[i+1] = PT(xL[i], c(r, k, 1))
    h = max(0, xL[i+1] - OptEscL)
    xL[i+1] = xL[i+1] - h
    objL = objL + h
  }  
  objL = objL + xL[N+1] #take all fish at eNd
  
## calculate the sum of harvest for perfect PT model   
  xPTp=rep(NA,N+1)
  xPTp[1]=k    
  objPTp=0    
  for(i in 1:N){
    xPTp[i+1] = PT(xPTp[i], c(r, k, phi))
    h = max(0, xPTp[i+1] - OptEscPTp)
    xPTp[i+1] = xPTp[i+1] - h
    objPTp = objPTp + h
  }  
  objPTp = objPTp + xPTp[N+1] #take all fish at eNd
  
```

The objective under perfect knowledge is `r objPTp` and using the PT function but with parameters estimated from the data `r objPT` and using the logistic escapement rule with parameters estimated from the data, `r objL`. This is despite the SSEs being `r optOutPT$value` and `r optOutL$value` respectively. Note the above objective is calculated using deterministic dynamics. I will soon include stochastic dynamics below.
